### 2024-07-03 causal-inference.qmd
20240703 - selection bias

Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied.

>>>>

Selection bias occurs when the selection of participants or their inclusion in analyses depends on the variables being studied.
For instance, if you are conducting a study on the extent to which sports drink consumption influences fantasy points, there would be selection bias if players are less likely to participate in the study if they score fewer fantasy points.

Now, consider a study in which you conduct a randomized controlled trial (RCT; i.e., an [experiment](#sec-experiment)) to evaluate the effect of a new medication on player performance.
You randomly assign some players to take the medication and other players to take a placebo.
Assume the new medication has side effects and leads many of the players who take it to drop out of the study.
This is an example of attrition bias (i.e., systematic attrition).
If you were to exclude these individuals from your analysis, it may make it appear that the medication led to better performance, because the players who experienced the side effect (and worse performance) dropped out of the study.
Hence, conducting an analysis that excludes these players from the analysis would involve selection bias.
